Combining formal and distributional models of predicate semantics

### 2016-03-14 papers/acl2016/paper.tex
Mostly-finished draft of the ACL 2016 paper
Former-commit-id: 5826274506c82738642a90457ebb0a27b8b24617

\documentclass[11pt]{article}
\usepackage{acl2016}
\usepackage{amsmath}
\usepackage{times}
\usepackage{latexsym}
\usepackage{color}
\usepackage{booktabs}
\usepackage{subfig}

@@

\begin{align*}
  P(c(e) = \true{}) &= \sigma ( \theta_c^T \phi_e ) \\
  P(r(e_1, e_2) = \true{}) &= \sigma ( \theta_r^T \phi_{(e_1, e_2)} )
\end{align*}

@@

In this paper, we use the system produced by K\&M, changing only the models of
predicate semantics, as described in \secref{method}.

@@

Subgraph feature extraction (SFE) is a technique introduced by Gardner and
Mitchell~\shortcite{gardner-2015-sfe} for generating feature matrices over node
pairs in a graph with labeled edges.  Given a pair of nodes in a graph, SFE
performs a search to characterize a subgraph around those two nodes, then
extracts various kinds of features from that subgraph, such as the set of edge
sequences that connect the two nodes.  \mattnote{Give an example graph with
features.}  Gardner and Mitchell used these features to perform link
prediction in the graph, using characteristics of the subgraph around two nodes
to determine if an edge of a particular type should exist between those nodes
(when the graph corresponds to a knowledge base like Freebase, this task is
also known as knowledge base completion). Gardner and Mitchell only used these
features to model formal KB relations, however.

@@

Some related work on including structured information in distributional representations: Manaal's
retrofitting semantic lexicons, Sebastian Riedel's ``Injecting Logical Background Knowledge into
Embeddings for Relation Extraction'', work on PRA and KBC more generally\ldots Also, the whole line
of work on knowledge base embeddings (RESCAL, TransE, etc.) is at least slightly relevant here, as
is PRA.  And all of the semantic parsing approaches that go from text to a pseudo-logical form,
then to an actual Freebase query (Kwiatkowski, Reddy's TACL paper, Liang?) \mattnote{Actually write
this section}.

@@

\subsection{Adding SFE features to Jayant's models}
\label{sec:formal-and-distributional}

The modification we make to K\&M's semantic parser is simple.  For each entity
and entity pair in the data we compute an SFE feature vector $\psi(e)$ and
$\psi(e_1, e_2)$.  For each category and relation, we select a subset of the
features to have associated weights $\omega_c$ and $\omega_r$.  We then add the
weights $\omega_c$ and $\omega_r$ as parameters to be learned for each
predicate, and we add the feature vectors $\psi(e)$ and $\psi(e_1, e_2)$ as
observed features for each entity and entity pair.  The augmented category and
relation probabilities are as follows:

\begin{align*}
  P(c(e) & = \true{}) & = & \sigma ( \theta_c^T \phi_e + \omega_c^T \psi_c(e)) \\
  P(r(e_1, e_2) & = \true{}) & = & \sigma ( \theta_r^T \phi_{(e_1, e_2)} + \omega_r^T \psi_r(e_1, e_2) )
\end{align*}
where we have added subscripts on the $\psi$ functions to indicate that each
predicate selects its own set of features from the SFE feature vector $\psi$.

Note that now there are three sets of parameters to be learned by the semantic
parser: (1) $\theta$, low-dimensional distributional vectors trained for each
predicate; (2) $\phi$, low-dimensional distributional vectors trained for each
entity and entity pair; and (3) $\omega$, weights associated with the selected
formal SFE features for each predicate.  All of these parameters are optimized
jointly, using the same ranking objective used by K\&M.

@@

PMI measures the strength of association between two variables $x$ and $y$; it
is computed as $\log\frac{p(x,y)}{p(x)p(y)}$.  In our case, we want to
associate predicates with SFE features.  However, our SFE features are for each
\emph{entity} and \emph{entity pair}, not each \emph{predicate}.  To associate
SFE features with predicates, we use the training data to find a mapping from
entities and entity pairs to predicates that they have been seen with in the
data.  For example, the phrase ``Italian architect Andrea Palladio'' is
considered a positive training example for the instantiated predicates
$\predicate{architect}(\entity{Andrea Palladio})$ and
$\predicate{architect\_N/N}(\entity{Italy}, \entity{Andrea Palladio})$.  We
then associate the feature vectors for \entity{Andrea Palladio} and
(\entity{Italy}, \entity{Andrea Palladio}) with the predicates
\predicate{architect} and \predicate{architect\_N/N}, respectively.  For each
predicate $\pi$ and feature $f$, we use these counts to calculate the
probabilities $p(\pi, f)$, $p(\pi)$ and $p(f)$.  After removing features with
counts below some threshold, we pick the $k$ features with the highest PMI
values for each predicate to use in our model.

@@

@@

\mattnote{I will add more detail to some of these examples; I ran out of
time...}

Our model does very well on most predicates that have clear Freebase
correlates: prime ministers, presidents, newspapers, companies, and so on.  The
query ``French newspaper \blank{}'' gets almost perfect average precision.

It does less well on predicates that do not match as closely with Freebase,
such as ``Microsoft head honcho \blank{}'', ``Patriots linebacker \blank{}'',
etc.  The combined model handily beats the distributional-only model in most of
these cases, but it does not have the nearly-perfect AP seen in the more crisp
queries above.  The ``head honcho'' example is particularly informative---this
is a phrase that essentially means ``leader'', and the features learned by the
formal model get some essential hints from Freebase, even though there is no
direct correlate.  The features capture the fact that entities that are
``honchos'' are people, and that they tend to be CEOs and film producers.

For ``Patriots linebacker \blank{}'', the issue is that the features for
\predicate{linebacker\_N/N} are able to capture which athletes play for which
teams, but they cannot distinguish well between what positions those athletes
play, so Patriots quarterbacks, wide receivers, and running backs are scored
highly for this query.  This could be fixed by using more complex features for
relation predicates, instead of just edge sequences, but that would explode the
feature space and make the feature selection process more difficult.  It could
also be fixed by better negative example selection during training.

There are also cases where a predicate has so many possible entities types in
Freebase that the it cannot learn models that are useful for all queries.  A
good example of this is ``\blank{} is a NASA mission''.

The SFE features we used cannot capture time very well.  ``17th century French
mathematician \blank{}''...

Because of the simple logical forms produced by the semantic parser we used, we
cannot model some complex phenomena, such as counting and other things.  This
is an orthogonal issue to the predicate semantics we're concerned with in this
paper, however.

Vice president Al Gore is modeled as $\predicate{vice}(\entity{Al Gore}) \land
\predicate{president}(\entity{Al Gore})$.  Clearly there is room for better
processing of multi-word expressions.

\predicate{head\_honcho\_N/N} was never seen in the training data, so the
phrase ``Microsoft head honcho \blank{}'' has to rely on knowing whether an
entity is a ``honcho'' and what entities are related to Microsoft by an unknown
predicate.  \predicate{honcho\_N/N} \emph{is} seen in the training data,
however; some kind of backoff to simpler phrases would be useful here.

Using connected entities in Freebase as additional candidates during inference
is helpful, but it is not very sophisticated.  It over-generates for a lot of
entities and predicates, and under-generates for others.  A mechanism for
tailoring a search over Freebase to a particular query would be useful, though
that would move the model more into the realm of reinforcement learning.

@@

\mattnote{TODO}

>>>>

\documentclass[11pt]{article}
\usepackage{acl2016}
\usepackage{amsmath}
\usepackage{times}
\usepackage{latexsym}
\usepackage{color}
\usepackage{booktabs}
\usepackage{subfig}
\usepackage{tikz}
\usetikzlibrary{bayesnet}

@@

\begin{align*}
  P(c(e)) &= \sigma ( \theta_c^T \phi_e ) \\
  P(r(e_1, e_2)) &= \sigma ( \theta_r^T \phi_{(e_1, e_2)} )
\end{align*}

@@

In this paper, we use the system produced by K\&M, changing only the models of
predicate semantics, as described in \secref{method}.  K\&M introduced several
variations of their parser; in this work, we use their query ranking objective,
as it had the best performance, and we evaluate the model they called
``Factorization'', without ensembling it with a clustering baseline.

@@

\begin{figure}
  {\center
  \begin{tikzpicture}[
    ->,
    shorten >=1pt,
    auto,
    node distance=3cm,
    thick,
    main node/.style={draw=none,fill=none}
    ]

    \node[main node] (1) {\entity{Palladio}};
    \node[main node] (2) [right=4cm of 1] {\entity{Italy}};
    \node[main node] (3) [below=1.7cm of 1] {\entity{architect}};
    \node[main node] (4) [below=1.7cm of 2] {\entity{country}};
    \node[main node] (5) [below right=.1cm and .5cm of 3] {\entity{Villa Capra}};

    \path[]
      (1) edge node [sloped, anchor=center, above] {\predicate{nationality}} (2)
      (1) edge node [sloped, anchor=center, above] {\predicate{type}} (3)
      (1) edge node [sloped, anchor=center, above] {\predicate{designed}} (5)
      (2) edge node [sloped, anchor=center, above] {\predicate{type}} (4)
      (5) edge node [sloped, anchor=center, above] {\predicate{located\_in}} (2);
  \end{tikzpicture}
  }
  \textbf{Features between \entity{Palladio} and \entity{Italy}}:\\
  \textless\predicate{nationality}\textgreater\\
  \textless\predicate{designed}, \predicate{located\_in}\textgreater
  \caption{A subset of the Freebase graph, and some example extracted features
  between \entity{Palladio} and \entity{Italy}.  The actual Freebase relations
  and entity identifiers used are modified here to aid readability.}
  \label{fig:sfe}
\end{figure}

Subgraph feature extraction (SFE) is a technique introduced by Gardner and
Mitchell~\shortcite{gardner-2015-sfe} for generating feature matrices over node
pairs in a graph with labeled edges.  Given a pair of nodes in a graph, SFE
performs a search to characterize a subgraph around those two nodes, then
extracts various kinds of features from that subgraph, such as the set of edge
sequences that connect the two nodes.  \figref{sfe} shows an example subgraph,
centered around \entity{Andrea Pallado} and \entity{Italy}, and features that
would be extracted for that entity pair.  Gardner and Mitchell used these
features to perform link prediction in the graph, using characteristics of the
subgraph around two nodes to determine if an edge of a particular type should
exist between those nodes.\footnote{When the graph corresponds to a knowledge
base like Freebase, this task is also known as knowledge base completion.}
Gardner and Mitchell only used these features to model formal KB relations,
however.

@@

There is a large body of other work that is closely related to what is
presented in this paper.  Two recent papers that also attempt to include some
form of structured knowledge into distributional representations are those by
Faruqui et al.~\shortcite{faruqui-2015-retrofitting-word-vectors}, and by
Rockt\"{a}schel et al.~\shortcite{rocktaschel-2015-logical-embeddings}.  A key
difference between their approaches and ours is that they use the structured
information only to modify distributional representations, while we include the
structured information directly in our models, training weights for structured
information jointly with distributional representations.  Also conceptually
similar is the Additive Relational Effects model for knowledge base completion
by Nickel et al.~\shortcite{nickel-2014-are}, which tries to factorize the
residual of a KB tensor that cannot be explained by observable features that
are somewhat similar to the formal features we use in this model.  But while
this model is conceptually similar, it is computationally quite different, and
would be rather difficult to use in the setting we explore in this work.

Our work draws a lot of inspiration from recent work on knowledge base
completion.  We are building on top of work in embedding the entities and
relations in a knowledge
base~\cite{riedel-2013-mf-universal-schema,nickel-2011-rescal,bordes-2013-transe},
as well as work on graph-based methods for reasoning with knowledge
bases~\cite{lao-2010-original-pra,gardner-2014-vector-space-pra,neelakantan-2015-rnn-kbc}.

Lastly, there is an extensive literature on building semantic parsers to answer
questions against a structured knowledge
base~\cite{zettlemoyer-2005-ccg,berant-2013-semantic-parsing-qa,%
kwiatkowski-2013-ontology-matching,krishnamurthy-2012-semantic-parsing}.  The
vast majority of this work is focused on matching natural language text to
statements in a specific ontology, with no hope of assigning meaning to
language that falls outside of that ontology.  Only very recently has there
been work attempting to learn broad coverage meaning representations in
semantic parsers.  In addition to the work by Krishnamurthy and Mitchell which
we extend in this paper, there was recently a nice piece of work by Choi et
al.~\shortcite{choi-2015-semantic-parsing-partial-ontologies}.  The aim of
their work is, like typical semantic parsers, to map text onto a single formal
statement, but they allow some of the predicates to be ``open'' predicates not
contained in the Freebase ontology.  A key difference between their work and
ours is that we score surface logical forms directly, using a weighted
combination of many Freebase queries, instead of trying to map them to a single
Freebase logical form.  Additionally, while their system might learn that
``front-runner'' should be modeled as an open predicate that has no
representation in Freebase, they have no method to assign any further meaning
to that predicate, while our models learn both distributional and formal
information about the meaning of ``front-runner''.

@@

\subsection{Formal features for predicate semantics}
\label{sec:formal-and-distributional}

We inject information from a formal knowledge base into our models of predicate
semantics simply by augmenting the learned distributional feature vectors with
formal features.

For each entity and entity pair in the data we compute an SFE feature vector
$\psi(e)$ and $\psi(e_1, e_2)$.  For entity pairs, the vector $\psi(e_1, e_2)$
is a sparse binary vector containing indicator features for all edge sequences
that connect nodes $e_1$ and $e_2$ in the Freebase graph, up to length 4.
These features are shown for the entity pair (\entity{Palladio},
\entity{Italy}) in the example graph in \figref{sfe}.  For entities, we extract
similar features to form the sparse binary vector $\psi(e)$.  These features
are of two types: (1) edge sequences that begin at node $e$ (up to length two),
and (2) edge sequences that begin at node $e$ along with the node at the end of
the sequence.  For example, all of the features listed in \figref{sfe} for the
entity pair (\entity{Palladio}, \entity{Italy}) are also valid features of type
(1) for the entity \entity{Palladio}, as well as the features
\textless\predicate{type}\textgreater and
\textless\predicate{designed}\textgreater.  Examples of of type (2) features
are \textless\predicate{type}\textgreater:\entity{architect}, and
\textless\predicate{nationality}\textgreater:\entity{Italy}.  Intuitively,
features of type (1) capture the fact that an entity participates in particular
kinds of relations, such as designing architectural structures, while features
of type (2) capture the way an entity is related to particular other entities.
Features of type (2) are very specific, and are most useful in practice when
the entity in the feature is a type node, such as with
\textless\predicate{type}\textgreater:\entity{architect}, though occasionally
features of this type involving other entities are also selected in our feature
selection process.

For each category and relation, we select a subset of the SFE features in the
vectors $\psi$ to have associated weights $\omega_c$ and $\omega_r$.  We then
add the weights $\omega_c$ and $\omega_r$ as parameters to be learned for each
predicate, and we add the feature vectors $\psi(e)$ and $\psi(e_1, e_2)$ as
observed features for each entity and entity pair.  The augmented category and
relation probabilities are as follows:

\begin{align*}
  P(c(e)) &= \sigma ( \theta_c^T \phi_e + \omega_c^T \psi_c(e)) \\
  P(r(e_1, e_2)) &= \sigma ( \theta_r^T \phi_{(e_1, e_2)} + \omega_r^T \psi_r(e_1, e_2) )
\end{align*}
where we have added subscripts on the $\psi$ functions to indicate that each
predicate selects its own set of features from the SFE feature vector $\psi$.

We pause here to note that these features obtained from SFE correspond exactly
to a limited set of horn clauses using the Freebase vocabulary, which in turn
can be seen as Freebase queries that define sets of entities or entity pairs.
By associating a number of these features with each predicate and assigning
them weights, what we are effectively doing is mapping language onto a weighted
combination of possible Freebase statements, instead of onto a single Freebase
query, as is typically done in the semantic parsing literature.  This is a
fundamentally new kind of predicate semantics that has not been used previously
in semantic parsers.

Note also that in our combined model there are now three sets of parameters to
be learned by the semantic parser: (1) $\theta$, low-dimensional distributional
vectors trained for each predicate; (2) $\phi$, low-dimensional distributional
vectors trained for each entity and entity pair; and (3) $\omega$, weights
associated with the selected formal SFE features for each predicate.  All of
these parameters are optimized jointly, using the same ranking objective used
by K\&M.

@@

PMI measures the strength of association between two variables $x$ and $y$; it
is computed as $\log\frac{p(x,y)}{p(x)p(y)}$.  In our case, we want to
associate predicates with SFE features.  However, our SFE features are for each
\emph{entity} and \emph{entity pair}, not each \emph{predicate}.  To associate
SFE features with predicates, we use the training data to find a mapping from
entities and entity pairs to predicates that they have been seen with in the
data.  For example, the phrase ``Italian architect Andrea Palladio'' is
considered a positive training example for the instantiated predicates
$\predicate{architect}(\entity{Andrea Palladio})$ and
$\predicate{architect\_N/N}(\entity{Italy}, \entity{Andrea Palladio})$.  We
then associate the feature vectors for \entity{Andrea Palladio} and
(\entity{Italy}, \entity{Andrea Palladio}) with the predicates
\predicate{architect} and \predicate{architect\_N/N}, respectively.  For each
predicate $\pi$ and feature $f$, we use these counts to calculate the
probabilities $p(\pi, f)$, $p(\pi)$ and $p(f)$.  After removing features with
counts below some threshold, we pick the $k$ features with the highest PMI
values for each predicate to use in our model.\footnote{We used $k=100$ in this
paper.}

@@

On both of these datasets, the difference between the combined model and the
distributional model is statistically significant (by a paired permutation
test, $p < 0.05$).  The difference between the combined model and the formal
model, and between the formal model and the distributional model, are not
statistically significant, as each method has certain kinds of queries that it
performs well on.  Only the combined model is able to consistently outperform
the distributional model on all kinds of queries.

@@

Here we discuss some of the successes and failures of our model.

When a predicate has a clear Freebase correlate (such a
\predicate{prime\_minister\_N/N}, \predicate{president\_of},
\predicate{newspaper}, \predicate{company}, etc.), our formal model is able to
capture the meaning very well.  The query ``French newspaper \blank{}'' has
nearly perfect average precision, for instance, as does ``Israeli prime
minister \blank{}''.  The distributional model does much less well here, and in
combination, the distributional component of the model occasionally lowers its
performance relative to the formal model.\footnote{This is why the difference
between the formal model and the combined model is not statistically
significant.}

The formal models also perform well on predicates that can be partially
expressed by Freebase, such as ``Microsoft head honcho \blank{}'', ``The United
States, \blank{}'s closest ally'', and ``Patriots linebacker \blank{}'', though
the models do not have the nearly-perfect AP that they do for the more crisp
queries above.  The ``head honcho'' example is particularly instructive---this
is a phrase that essentially means ``leader'', and the features learned by the
formal model get some essential hints from Freebase, even though there is no
direct correlate.  The features capture the fact that entities that are
``honchos'' are people, and that they tend to be CEOs and film producers.

For ``Patriots linebacker \blank{}'', the reason that this can only be
partially expressed in terms of Freebase queries is that the features for
\predicate{linebacker\_N/N} are able to capture which athletes play for which
teams, but they cannot distinguish well between what positions those athletes
play.  This leads to Patriots quarterbacks, wide receivers, and running backs
being scored highly for this query.  This issue could be fixed by using more
complex features for relation predicates, instead of just edge sequences,
though that would drammatically increase the size of the feature space and make
the feature selection process much more difficult.  It could also potentially
be mitigated by better negative example selection during training.

There are also cases where a predicate has so many possible entity types in
Freebase (or no close correlate at all) that the PMI feature selection process
does not find features that are useful for all queries.  A good example of this
is ``\blank{} is a NASA mission''; in this case, the distributional model
outperforms the formal model, because it has learned associations between
particular NASA missions and NASA.

While the new logical forms we introduced in \secref{better-lfs} improved the
performance of the models, there is still a lot of room for further
improvement.  For instance, ``Vice president Al Gore'' is modeled as
$\predicate{vice}(\entity{Al Gore}) \land \predicate{president}(\entity{Al
Gore})$.  Better processing of multi-word expressions would allow for more
crisp models of the predicates, so the model for \predicate{president} could
have more specific features.  Also, the predicate \predicate{head\_honcho\_N/N}
was never seen in the training data, so an \predicate{unknown} predicate was
used to score the relation in the phrase ``Microsoft head honcho \blank{}''.
This means the model had to rely solely on knowing which entities are
``honchos'' and which entities are related to Microsoft in some unknown manner.
The predicate \predicate{honcho\_N/N} \emph{is} seen in the training data,
however; some kind of backoff to simpler phrases for unseen predicates would be
useful here.

Finally, using connected entities in Freebase as additional candidates during
inference is helpful, but it is not very sophisticated.  It over-generates for
a lot of entities and predicates, and under-generates for others.  A mechanism
for tailoring a search over Freebase to a particular query would be useful,
though it would add a lot of additional complexity, possibly requiring methods
from reinforcement learning to resolve.

@@

In this paper, we have considered the problem of assigning meaning to
predicates (both categories and relations) in natural language, using both
distributional information and information from a formal knowledge base.  To do
this, we developed a model that can represent words as a weighted combination
of formal statements, and trained that model jointly with a distributional
model of word meaning.  This joint model achieved a relative gain of 62\% in
mean average precision over the distributional model alone.  We also introduced
a better mapping from surface text to logical forms, and a simple method for
using Freebase to find candidate entities during inference.  Taken together,
the methods introduced in this paper improved mean average precision on a
fill-in-the-blank natural language query task from .163 to .370, a 127\%
relative improvement.
